---
title: "Breeds impact on adoption"
author: "Matthew Wilcox, Randy Otoo, Anthony Barrios, Arianne PozziBarba"
output: pdf_document
---

# Libraries

```{r}
library(tidyverse)
library(knitr)
library(kableExtra)

```
# Import and Clean Data

```{r}
animal_intake.df <- read_csv("Louisville_Metro_ky_-_Animal_Service_Intake_and_Outcome.csv")

```

```{r}
dog_intake.df <- animal_intake.df %>% filter(animaltype == "DOG") %>% 
  mutate(indate = str_remove(indate, " .*"),
         outdate = str_remove(outdate, " .*"),
         indate = as.Date(indate, format = "%Y/%m/%d"),
         outdate = as.Date(outdate, format = "%Y/%m/%d"),
         timeHeld = outdate - indate
         )%>%
  filter(intype == "STRAY", 
         outtype != "DISPOSAL") %>%
  select(outtype, sex, bites, petsize, color, breed, timeHeld)


```

# Probability

```{r}
pitbull_adopt_euth <- dog_intake.df %>%
  select(outtype, breed) %>%
  filter(
    outtype == "ADOPTION" | outtype == "EUTH"
  ) %>%
  mutate(breed = case_when(
    grepl("PIT BULL", breed) ~ "Is_Pitbull",
    TRUE ~ "Not_Pitbull"
  ))%>% group_by(outtype, breed) %>%
  summarise(n = n()) %>%
  pivot_wider(names_from = breed, values_from = n) 
pitbull_adopt_euth_display <- pitbull_adopt_euth %>%
  ungroup() %>%
  add_row(outtype = "TOTAL",
          Is_Pitbull = sum(pitbull_adopt_euth$Is_Pitbull), 
          Not_Pitbull = sum(pitbull_adopt_euth$Not_Pitbull)
          ) %>%
  mutate( Total = Is_Pitbull + Not_Pitbull) %>%
  column_to_rownames(., var = "outtype") 
pitbull_adopt_euth <- pitbull_adopt_euth %>%
  column_to_rownames(., var = "outtype") 



pitbull_adopt_euth_display %>% kbl(caption = "Frequency Table of Pitbull and Adoption") %>%  kable_classic_2(full_width = F)


```


```{r}
total <- sum(pitbull_adopt_euth)

# Probability of a (Prob is pitbull)
a<- sum(pitbull_adopt_euth[1])/total

# Probaility of (Adoption)

b<- sum(pitbull_adopt_euth[1,])/total

# Probability (b | a) (Prob Adopted given pitbull)
bGivena<- pitbull_adopt_euth[1,1]/sum(pitbull_adopt_euth[1])
 
# Probability (b | ¬a)
bGivenNota<- pitbull_adopt_euth[1,2]/sum(pitbull_adopt_euth[2])

# Probibility of Ands
aANDb <- a * bGivena
aANDnotb <- a * (1-bGivena)
notaANDb<- (1-a) * bGivenNota
notaANDnotb <- (1-a) *(1-bGivenNota)


#Baysian
aGivenb <- aANDb/b
aGivenNotb <- aANDnotb/(1-b)

```

```{r}
prob_type_col <- c("P(Is a Pitbull)", 
                   "P(Not a Pitbull)", 
                   "P(Adopted)", 
                   "P(Euthanized)", 
                   "P(Adopted | Is a Pitbull)",
                   "P(Euthanized | Is a Pitbull)",
                   "P(Adopted | Not A Pitbull)",
                   "P(Euthanized | Not a Pitbull)",
                   "P(Pitbull & Adopted)",
                   "P(Pitbull & Euthanized)",
                   "P(Not a Pitbull & Adopted)",
                   "P(Not a Pitbull & Euthanized)",
                   "p(Pitbull | Adopted)",
                   "P(Not a Pitbull | Adopted)",
                   "P(Pitbull | Euthanized)",
                   "P(Not a Pitbull | Euthanized)"
                   )
prob_col<- c(paste(round(a*100,2),"%", sep = ""), #P(pitbull)
             paste(round((1-a)*100,2),"%", sep = ""), #P(not pitbull)
             paste(round((b)*100,2),"%", sep = ""), #P(Adopted)
             paste(round((1-b)*100,2),"%", sep = ""),#P(euthanized)
             paste(round((bGivena)*100,2),"%", sep = ""), #P(Adopted|pitbull)
             paste(round((1-bGivena)*100,2),"%", sep = ""), #P(Euthanized|Pitbull)
             paste(round((bGivenNota)*100,2),"%", sep = ""), #P(Adopted|not pit)
             paste(round((1-bGivenNota)*100,2),"%", sep = ""), #P(Euthanized|notpit)
             paste(round((aANDb)*100,2),"%", sep = ""), #P(P&A)
             paste(round((aANDnotb)*100,2),"%", sep = ""), #P(P&E)
             paste(round((notaANDb)*100,2),"%", sep = ""), #P(p'&A)
             paste(round((notaANDnotb)*100,2),"%", sep = ""), #P(p1&E)
             paste(round((aGivenb)*100,2),"%", sep = ""), # p(Pitbull | Adopted)
             paste(round((1-aGivenb)*100,2),"%", sep = ""), # P(Not a Pitbull | Adopted)
             paste(round((aGivenNotb)*100,2),"%", sep = ""), # P(Pitbull | Euthanized)
             paste(round((1-aGivenNotb)*100,2),"%", sep = "") #P(Not a Pitbull | Euthanized)
             )
prob_df <- data.frame(Condition = prob_type_col, Probability = prob_col)
prob_df %>% kbl()  %>%
  kable_styling()
```


```{r}
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("Rgraphviz")
library("Rgraphviz")
```

```{r}
total <- sum(pitbull_adopt_euth)

# Probability of a (Prob is pitbull)
a<- sum(pitbull_adopt_euth[1])/total
 
# Probability (b | a) (Prob Adopted given pitbull)
bGivena<- pitbull_adopt_euth[1,1]/sum(pitbull_adopt_euth[1])
 
# Probability (b | ¬a)
bGivenNota<- pitbull_adopt_euth[1,2]/sum(pitbull_adopt_euth[2])


###################### Everything below here will be calculated
 
# Calculate the rest of the values based upon the 3 variables above
notbGivena<-1-bGivena
notA<-1-a
notbGivenNota<-1-bGivenNota
 
#Joint Probabilities of a and B, a and notb, nota and b, nota and notb
aANDb<-a*bGivena
aANDnotb<-a*notbGivena
notaANDb <- notA*bGivenNota
notaANDnotb <- notA*notbGivenNota
 
# Probability of B
b<- aANDb + notaANDb
notB <- 1-b
 
# Bayes theorum - probabiliyt of A | B
# (a | b) = Prob (a AND b) / prob (b)
aGivenb <- aANDb / b
 
# These are the labels of the nodes on the graph
# To signify "Not A" - we use A' or A prime 
 
node1<-"P"
node2<-"Is Pitbull"
node3<-"Not Pitbull"
node4<-"IS Pitbull & Adopted"
node5<-"Is Pitbull & Euthanized'"
node6<-"Not Pitbull & Adopted"
node7<-"Not Pitbull & Euthanized"
nodeNames<-c(node1,node2,node3,node4, node5,node6, node7)
 
rEG <- new("graphNEL", nodes=nodeNames, edgemode="directed")
#Erase any existing plots
# dev.off()
 
# Draw the "lines" or "branches" of the probability Tree
rEG <- addEdge(nodeNames[1], nodeNames[2], rEG, 1)
rEG <- addEdge(nodeNames[1], nodeNames[3], rEG, 1)
rEG <- addEdge(nodeNames[2], nodeNames[4], rEG, 1)
rEG <- addEdge(nodeNames[2], nodeNames[5], rEG, 1)
rEG <- addEdge(nodeNames[3], nodeNames[6], rEG, 1)
rEG <- addEdge(nodeNames[3], nodeNames[7], rEG, 10)
 
eAttrs <- list()
 
q<-edgeNames(rEG)
 
# Add the probability values to the the branch lines
 
eAttrs$label <- c(toString(round(a,4)),toString(round(notA,4)),
 toString(round(bGivena,4)), toString(round(notbGivena,4)),
 toString(round(bGivenNota,4)), toString(round(notbGivenNota,4)))
names(eAttrs$label) <- c(q[1],q[2], q[3], q[4], q[5], q[6])
edgeAttrs<-eAttrs
 
# Set the color, etc, of the tree
attributes<-list(node=list(label="foo", fillcolor="lightgreen", fontsize="15"),
 edge=list(color="red"),graph=list(rankdir="LR"))
 
#Plot the probability tree using Rgraphvis
plot(rEG, edgeAttrs=eAttrs, attrs=attributes)
nodes(rEG)
edges(rEG)
 
#Add the probability values to the leaves of A&B, A&B', A'&B, A'&B'
text(500,420,round(aANDb,4), cex=.8)
 
text(500,280,round(aANDnotb,4),cex=.8)
 
text(500,160,round(notaANDb,4),cex=.8)
 
text(500,30,round(notaANDnotb,4),cex=.8)
 
text(340,440,"(Adopted | Is Pitbull)",cex=.8)
 
text(340,230,"(Euthanized | Is Pitbull)",cex=.8)
 
#Write a table in the lower left of the probablites of A and B
text(80,50,paste("P(A):",round(a,4)),cex=.9, col="darkgreen")
text(80,20,paste("P(A'):",round(notA,4)),cex=.9, col="darkgreen")
 
text(160,50,paste("P(B):",round(b,digits=4)),cex=.9)
text(160,20,paste("P(B'):",round(notB, 4)),cex=.9)
 
text(80,420,paste("P(A|B): ",round(aGivenb,digits=4)),cex=.9,col="blue")

```  

https://www.harrysurden.com/wordpress/archives/292 refrenced to help build the tree diagram. 





# Hypthosesis 
Null Hypothesis (H0): The populations average length of stay for Pit Bulls is equal to the general populations average length of stay.
Alternative Hypothesis (HA): The populations average length of stay for Pit Bulls is different from the general population's average length of stay

```{r}

dog_intake.df <- dog_intake.df %>%
  mutate(Pitbull = ifelse(grepl("PIT BULL", breed), "Pitbull", "Other Breed"))
pitbull_data <- dog_intake.df %>%
  filter(Pitbull == "Pitbull") %>%
  select(timeHeld)

other_breeds_data <- dog_intake.df %>%
  filter(Pitbull == "Other Breed") %>%
  select(timeHeld)
t_test_pitbull_vs_other_breeds <- t.test(pitbull_data$timeHeld, other_breeds_data$timeHeld)
t_test_pitbull_vs_other_breeds

```
We reject the null hypothesis because we have a small p-value (p < 0.05) This means there's significant evidence to suggest that the average stay duration of Pit bulls differs from the average stay duration of other dog breeds in the Louisville metro area.

```{r}
#Confidence interval
n <- sum(!is.na(dog_intake.df$timeHeld)) 
mean_stay <- mean(dog_intake.df$timeHeld, na.rm = TRUE) 
sd_stay <- sd(dog_intake.df$timeHeld, na.rm = TRUE)  

lower_ci <- mean_stay - qt(0.975, df = n - 1) * (sd_stay / sqrt(n))
upper_ci <- mean_stay + qt(0.975, df = n - 1) * (sd_stay / sqrt(n))

lower_ci 
upper_ci  

```
